NDNS multilevel LCA CW2CB2 or CW2CB3

### 2018-09-04 paper/Chapters/Chapter2.tex
right before submission checked once more

Participants were required to keep a diary that record everything eaten or drunk over four consecutive days. Interviewers undertook three visits with each participant. At the first visit, the interviewer explained the method followed a protocol, taking participants through the sections in the diary including how to describe details of food and drink and portion size and an example day. The second visit was a brief one to check for compliance, and answer potential questions from the participants. Interviewers also reviewed the diary to identify and edited possible omissions and missing detail in the food diary. The third visit was to collect the diary and again review and edit possible omissions.

In the diary, portion sizes of food or drink eaten were asked to be recorded in household measures (e.g. one tablespoon of beans, one Kit Kat finger-size), or for packaged foods to note the weight indicated on the packet. Homemade dishes were also described in the diary about their ingredients, quantities, along with a brief description of the cooking procedure and how much of the dish were consumed. For each eating occasion, in addition to the details of what and how much was eaten, participants were also asked to record: when was it, where they were, and who they were eating with. An example, used as guidance for participants, of a food diary for one day is shown in \textbf{Appendix \ref{AppendixE}}.\vspace{-0.3cm}

@@

\begin{itemize}
	\item Randomly select PSUs from the Postcode Address File; 
	\item Randomly select 27 or 28 addresses in that postcode area; 
	\item Randomly select one household at that address for interviews; 
	\item Selected addresses where children resided were randomly allocated to one of two groups to determine whether an adult (aged 19 years or older) and a child (aged 1.5 to 18 years) or a child only, were selected for interviews.
\end{itemize}
\vspace{-0.6cm}
%-----------------------------------
%	SUBSECTION 2
%-----------------------------------
\subsection{Response rates}\vspace{-0.3cm}

@@

Blood samples were stored at 4 $^\circ$C, and sent directly by post to the Department of Haematology and Department of Clinical Biochemistry and Immunology, Addenbrooke's Hospital, Cambridge within two hours of their collection. Serum samples were collected by centrifugation of the coagulated blood sample. Serum total, High-Density Lipoprotein (HDL) and Low-Density Lipoprotein (LDL) cholesterol, triglycerides (TG), fasting blood glucose, haemoglobin HbHbA1C were measured. HbA1C value of 6.5\% was used as the cut off point for diagnosing diabetes.

@@

\begin{itemize}
	\item $ U_{ijk} $ represents the response of carbohydrate eating (one of the following: not eating any food, $< 50\%$ of the energy, or $\geqslant 50\%$ of the energy) on the $i$th day of the survey ($i \in (1,2,3,4)$) in $j$th individual at the $k$th time slot of the day ($k \in (1, 2, 3, \cdots, 7)$);
	\item $G_{ij}$ denotes the latent class membership for $j$th individuals on the $i$th day of the survey, the total number of day-level latent class is $T$;
%	\item A specific latent class is referred to as $t$, and the total number of level 1 latent classes is denoted by $T$;
	\item $P(\mathbf{U_{ijk} = s_k}|C_{ij} = t)$ is the probability of a specific response pattern, conditional on membership in latent class $t$.
\end{itemize}

@@

\begin{itemize}
	\item First, we ignored the multilevel structure of the data and estimated a series of traditional LC models to determine the number of classes at the observational day-level;
	\item Next, a series of MLCA models were fitted to account for the multilevel structure of the data. In these models, the number of day-level classes was based on the best fitting LCA model from the first step, and the LCA model at the individual-level was estimated to identify the number of individual-level latent classes;
	\item Last, once the number of individual-level latent classes was defined based on the previous stage, the number of day-level classes was modified (one class lower and one class higher than in the second step) to investigate the effect of changing level 1 classes and to confirm the best fitting model.
\end{itemize}

@@

Day-level latent classes identified by the first step of MLCA were tabulated according to the day of the week and also whether the diary was recorded during weekends or not. A contingency table giving the frequency of responses across the 7-time slots of the survey days was produced. Descriptive statistics for the dietary day-level recordings according to the latent class memberships were presented. Pearson $\chi^2$ test was used to assess evidence fora difference in the distribution of categorical variables. One-way Analysis of Variance (ANOVA) was used to compare the means across the multiple groups for continuous variables.

@@

Survey-designed logistic regression models (for hypertension), and linear regression models (for WC, BMI), were used to test for associations between latent classes of carbohydrate eating patterns and hypertension, BMI, and WC, in the NDNS RP sample, separately. Since diabetic participants might or might not modify their carbohydrate eating habits, we also fitted all the above-mentioned regression models restricted to those without diabetes.

@@

\begin{enumerate}
	\item The crude association between the carbohydrate eating groups and the outcomes was first examined. 
	\item Potential confounders of the association between carbohydrate eating groups (exposure) and the outcomes were selected depending on the descriptive statistical analyses conducted above, i.e. those associated with both the exposure and the outcome and also not on the causal pathway were selected as potential confounders. Covariates that are strongly related with the outcomes but not associate with carbohydrate eating groups may reduce the standard errors and improve the precisions, so they were also considered in the linear regression models. 
	\item Confounding or interaction effects from each of the potential factors were checked one by one. Interaction effects were tested using the adjusted Wald test for whether the regression coefficients of the interaction terms are simultaneously equal to zero.  
	\item A preliminary model that includes all of the variables suggested to be confounders in the previous step was established. 
	\item The remaining variables were added to the preliminary model one by one to see if any of them may be a confounder in the presence of the other covariates. 
	\item For logistic regression models (hypertension) under the survey data, goodness-of-fit was assessed using \textit{F}-adjusted mean residual test \parencite{archer2001goodness,archer2007goodness} which is implemented in Stata using command \code{svylogitgof} \parencite{archer2006goodness}. Other diagnostics for regular logistic regression models, such as estimating the pseudo-R\textsuperscript{2},  AIC or BIC, checking the standardised Pearson residuals, or covariate pattern residuals are currently not available for weighted survey data. 
	\item For linear regression models (WC, BMI), the assumption of independent observations is violated because of the clustered structure of the data. Therefore, inference from such models is based on design rather than the model.  
	General checking such as QQ plots of the residuals (normality), plotting the residuals against fitted values (constant variance) are unneeded (also not available). Outliers, leverage, and Cook's distance cannot be checked either. However, participants with extreme weightings (if any) were checked by removing them and refitting the models as a sensitivity analysis.
	\item Since under complex survey designed data, the sampling-weighted least squares are not maximum likelihood, it would not be possible to compare models using likelihood ratio test. Instead, adjusted Wald tests with $p < 0.05$ were used as criteria for variable inclusion in the final model. Another Stata command \code{linktest} was used to decide whether quadratic and cubic terms of continuous variables were necessary for improving the fitting of the model \parencite{pregibon1980goodness}. 
\end{enumerate}

>>>>

Participants were required to keep a diary that record everything eaten or drunk over four consecutive days. Interviewers undertook three visits with each participant. At the first visit, the interviewer explained the method followed a protocol, taking participants through the sections in the diary including how to describe details of food and drink and portion size and an example day. The second visit was a brief one to check for compliance, and answer potential questions from the participants. Interviewers also reviewed the diary to identify and edited possible omissions and missing details in the food diary. The third visit was to collect the diary and again review and edit possible omissions.

In the diary, portion sizes of food or drink consumed were asked to be recorded in household measures (e.g. one tablespoon of beans, one Kit Kat finger-size), or for packaged foods to note the weight indicated on the packet. Homemade dishes were also described in the diary about their ingredients, quantities, along with a brief description of the cooking procedure and how much of the dish were consumed. For each eating occasion, in addition to the details of what and how much was eaten, participants were also asked to record: when was it, where they were, and who they were eating with. An example, used as guidance for participants, of a food diary for one day is shown in \textbf{Appendix \ref{AppendixE}}.\vspace{-0.3cm}

@@

\begin{itemize}
	\item Randomly select PSUs from the Postcode Address File; 
	\item Randomly select 27 or 28 addresses in that postcode area; 
	\item Randomly select one household at that address for interview; 
	\item Selected addresses where children resided were randomly allocated to one of two groups to determine whether an adult (aged 19 years or older) and a child (aged 1.5 to 18 years) or a child only, were selected for interviews.
\end{itemize}
\vspace{-0.6cm}
%-----------------------------------
%	SUBSECTION 2
%-----------------------------------
\subsection{Response rates}\vspace{-0.3cm}

@@

Blood samples were stored at 4 $^\circ$C, and sent directly by post to the Department of Haematology and Department of Clinical Biochemistry and Immunology, Addenbrooke's Hospital, Cambridge within two hours of their collection. Serum samples were collected by centrifugation of the coagulated blood sample. Serum total, High-Density Lipoprotein (HDL) and Low-Density Lipoprotein (LDL) cholesterol, triglycerides (TG), fasting blood glucose, haemoglobin(Hb) A1C were measured. HbA1C value of 6.5\% was used as the cut off point for diagnosing diabetes.

@@

\begin{itemize}
	\item $ U_{ijk} $ represents the response of carbohydrate eating (one of the following: not eating any food, $< 50\%$ of the energy, or $\geqslant 50\%$ of the energy) on the $i$th day of the survey ($i \in (1,2,3,4)$) in $j$th individual at the $k$th time slot of the day ($k \in (1, 2, 3, \cdots, 7)$);
	\item $G_{ij}$ denotes the latent class membership for $j$th individuals on the $i$th day of the survey, the total number of day-level latent class is $T$;
%	\item A specific latent class is referred to as $t$, and the total number of level 1 latent classes is denoted by $T$;
	\item $P(\mathbf{U_{ijk} = s_k}|G_{ij} = t)$ is the probability of a specific response pattern, conditional on membership in latent class $t$.
\end{itemize}

@@

\begin{itemize}
	\item First, we ignored the multilevel structure of the data and estimated a series of traditional LCA models to determine the number of classes at the observational day-level;
	\item Next, a series of MLCA models were fitted to account for the multilevel structure of the data. In these models, the number of day-level classes was based on the best fitting LCA model from the first step, and the LCA model at the individual-level was estimated to identify the number of individual-level latent classes;
	\item Last, once the number of individual-level latent classes was defined based on the previous stage, the number of day-level classes was modified (one class lower and one class higher than in the second step) to investigate the effect of changing level 1 classes and to confirm the best fitting model.
\end{itemize}

@@

Day-level latent classes identified by the first step of MLCA were tabulated according to the day of the week and also whether the diary was recorded during weekends or not. A contingency table giving the frequency of responses across the 7-time slots of the survey days was produced. Descriptive statistics for the dietary day-level recordings according to the latent class memberships were presented. Pearson $\chi^2$ test was used to assess evidence for a difference in the distribution of categorical variables. One-way Analysis of Variance (ANOVA) was used to compare the means across the multiple groups for continuous variables.

@@

Survey-designed logistic regression models (for hypertension), and survey-designed linear regression models (for WC, BMI), were used to test for associations between latent classes of carbohydrate eating patterns and hypertension, BMI, and WC, in the NDNS RP sample, separately. Since diabetic participants might or might not modify their carbohydrate eating habits, we also fitted all the above-mentioned regression models restricted to those without diabetes.

@@

\begin{enumerate}
	\item The crude association between the carbohydrate eating groups and the outcomes was first examined. 
	\item Potential confounders of the association between carbohydrate eating groups (exposure) and the outcomes were selected depending on the descriptive statistical analyses conducted above, i.e. those associated with both the exposure and the outcome and also not on the causal pathway were selected as potential confounders. Covariates that are strongly related with the outcomes but not associate with carbohydrate eating groups may reduce the standard errors and improve the precisions, so they were also considered in the survey-designed linear regression models. 
	\item Confounding or interaction effects from each of the potential factors were checked one by one. Interaction effects were tested using the adjusted Wald test for whether the regression coefficients of the interaction terms are simultaneously equal to zero.  
	\item A preliminary model that includes all of the variables suggested to be confounders in the previous step was established. 
	\item The remaining variables were added to the preliminary model one by one to see if any of them may be a confounder in the presence of the other covariates. 
	\item For logistic regression models (hypertension) under the survey data, goodness-of-fit was assessed using \textit{F}-adjusted mean residual test \parencite{archer2001goodness,archer2007goodness} which is implemented in Stata using command \code{svylogitgof} \parencite{archer2006goodness}. Other diagnostics for regular logistic regression models, such as estimating the pseudo-R\textsuperscript{2},  AIC or BIC, checking the standardised Pearson residuals, or covariate pattern residuals are currently not available for weighted survey data. 
	\item For linear regression models (WC, BMI), the assumption of independent observations is violated because of the clustered structure of the data. Therefore, inference from such models is based on design rather than the model.  
	General checking such as QQ plots of the residuals (normality), plotting the residuals against fitted values (constant variance) are unneeded (also not available). Outliers, leverage, and Cook's distance cannot be checked either. However, participants with extreme weightings (if any) were checked by removing them and refitting the models as a sensitivity analysis.
	\item Since under complex survey designed data, the sampling-weighted least squares are not maximum likelihood, it would not be possible to compare models using likelihood ratio test. Instead, adjusted Wald tests with $p < 0.05$ were used as criteria for variable inclusion in the final model. Another Stata command \code{linktest} was used to decide whether quadratic and cubic terms of continuous variables were necessary for improving the fitting of the model \parencite{pregibon1980goodness}. 
\end{enumerate}

### 2018-09-04 paper/Chapters/Chapter3.tex
pdf prepared to be submitted
right before submission checked once more

	\caption{\label{tab:daylevel}Day level latent class solution for three classes LCA model. (No individual level model)}\vspace{-0.3cm}
	\centering
	\fontsize{9}{11}\selectfont
	\begin{tabular}[t]{llccccc}
		\hiderowcolors
		\toprule
		\textbf{Time slots of} & \textbf{Responses to} & \multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \textbf{\Centerstack{Class 1 days\\(39.5\%)}} & \textbf{\Centerstack{Class 2 days\\(20.4\%)}} & \textbf{\Centerstack{Class 3 days\\(40.1\%)}} \\
		\cmidrule(l{2pt}r{2pt}){5-5} \cmidrule(l{2pt}r{2pt}){6-6} \cmidrule(l{2pt}r{2pt}){7-7}
		 \textbf{the day} &  \textbf{carbohydrate intake} & $n$ & (\%) & \textbf{\Centerstack{High perc-\\entage carb}} & \textbf{\Centerstack{Low perc-\\entage carb}} & \textbf{\Centerstack{Regular\\meals}}\\
		\midrule
		\showrowcolors
		6 am – 9 am &  &  &  &  &  & \\
		& Not eating any food & 7655 & 31.2 & 0.129 & 0.450 & 0.320\\
		& Carbohydrate $<$ 50\%\textsuperscript{*} & 4500 & 18.4 & 0.130 & 0.267 & 0.128\\
		& Carbohydrate $\geqslant$ 50\%\textsuperscript{\dag} & 12328 & 50.4 & 0.741 & 0.283 & 0.552\\
		9 am – 12 noon &  &  &  &  &  & \\
		& Not eating any food & 5447 & 22.2 & 0.237 & 0.079 & 0.401\\
		& Carbohydrate $<$ 50\% & 7227 & 29.5 & 0.158 & 0.492 & 0.173\\
		& Carbohydrate $\geqslant$ 50\% & 11809 & 48.2 & 0.605 & 0.429 & 0.426\\
		12 noon – 2 pm &  &  &  &  &  & \\
		& Not eating any food & 4783 & 19.5 & 0.156 & 0.356 & 0.019\\
		& Carbohydrate $<$ 50\% & 11112 & 45.4 & 0.405 & 0.413 & 0.560\\
		& Carbohydrate $\geqslant$ 50\% & 8588 & 35.1 & 0.439 & 0.231 & 0.421\\
		2 pm – 5 pm &  &  &  &  &  & \\
		& Not eating any food & 6926 & 28.3 & 0.130 & 0.123 & 0.659\\
		& Carbohydrate $<$ 50\% & 8277 & 33.8 & 0.249 & 0.602 & 0.076\\
		& Carbohydrate $\geqslant$ 50\% & 9280 & 37.9 & 0.621 & 0.276 & 0.266\\
		5 pm – 8 pm &  &  &  &  &  & \\
		& Not eating any food & 3043 & 12.4 & 0.114 & 0.199 & 0.034\\
		& Carbohydrate $<$ 50\% & 14240 & 58.2 & 0.516 & 0.590 & 0.639\\
		& Carbohydrate $\geqslant$ 50\% & 7200 & 29.4 & 0.370 & 0.211 & 0.328\\
		8 pm – 10 pm &  &  &  &  &  & \\
		& Not eating any food & 8722 & 35.6 & 0.322 & 0.291 & 0.480\\
		& Carbohydrate $<$ 50\% & 8898 & 36.3 & 0.266 & 0.551 & 0.212\\
		& Carbohydrate $\geqslant$ 50\% & 6863 & 28.0 & 0.412 & 0.158 & 0.308\\
		10 pm – 6 am &  &  &  &  &  & \\
		& Not eating any food & 16295 & 66.6 & 0.680 & 0.590 & 0.751\\
		& Carbohydrate $<$ 50\% & 4144 & 16.9 & 0.074 & 0.294 & 0.101\\
		& Carbohydrate $\geqslant$ 50\% & 4044 & 16.5 & 0.246 & 0.115 & 0.148\\
		\bottomrule
		\multicolumn{7}{l}{{\scriptsize \textit{Note: }}}\\
		\multicolumn{7}{l}{\scriptsize \textbf{Abbreviation:} LCA, latent class analysis, carb is short for carbohydrates.}\\
		\multicolumn{7}{l}{{\scriptsize \textsuperscript{*} Carbohydrate $<$ 50\% indicates that within the time slot, carbohydrate contributed $<$ 50\% total energy intake.}}\\
		\multicolumn{7}{l}{{\scriptsize \textsuperscript{\dag} Carbohydrate $\geqslant$ 50\% indicates that within the time slot, carbohydrate contributed $\geqslant$ 50\% total energy intake.}}\\
		\multicolumn{7}{l}{}\\ 
	\end{tabular}
\end{table}

@@

From the results of blood tests, 6.9\% of low carbohydrate eaters were found to be diabetic (diagnosed by HbA1C > 6.5\%), while the percentages of diabetes in the moderate and high carbohydrate eaters were 3.5\%, and 4.1\% (\textit{p} < 0.011), respectively. Although there was some evidence (\textit{p} = 0.027) that fasting blood glucose level may be slightly higher in non-diabetic low carbohydrate eaters, the geometric mean for HbA1C was probably lower (\textit{p} = 0.010) in moderate carbohydrate eaters (5.43, 95\%CI: 5.39, 5.47). Total cholesterol, HDL, and LDL were all lower in the moderate carbohydrate eaters, while no evidence of any difference of TG was found across three types of carbohydrate eaters.

@@

%		\multicolumn{5}{l}{\textit{Note: }}\\

@@

		\multicolumn{5}{l}{{\scriptsize BMI body mass index; WC, waist circumference; HbA1C, haemoglobin HbA1C; DM, diabetes mellitus; TC, total cholesterol, }}\\
		\multicolumn{5}{l}{{\scriptsize HDL, high density lipoproteins; LDL, low density lipoproteins; TG, triglycerides.}} \\
		\multicolumn{5}{l}{{\scriptsize Glucose, HbA1C, TC, HDL, LDL, TG, and physical activity were expressed in geometric means (95\% CI) because }} \\
		\multicolumn{5}{l}{ the data were positively skewed.} \\

>>>>

	\caption{\label{tab:daylevel}Day level latent class solution for three classes LCA model. (No individual level model)}\vspace{-0.3cm}
	\centering
	\fontsize{9}{11}\selectfont
	\begin{tabular}[t]{llccccc}
		\hiderowcolors
		\toprule
		\textbf{Time slots of} & \textbf{Responses to} & \multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \textbf{\Centerstack{Class 1 days\\(39.5\%)}} & \textbf{\Centerstack{Class 2 days\\(20.4\%)}} & \textbf{\Centerstack{Class 3 days\\(40.1\%)}} \\
		\cmidrule(l{2pt}r{2pt}){5-5} \cmidrule(l{2pt}r{2pt}){6-6} \cmidrule(l{2pt}r{2pt}){7-7}
		 \textbf{the day} &  \textbf{carbohydrate intake} & $n$ & \% & \textbf{\Centerstack{High perc-\\entage carb}} & \textbf{\Centerstack{Low perc-\\entage carb}} & \textbf{\Centerstack{Regular\\meals}}\\
		\midrule
		\showrowcolors
		6 am – 9 am &  &  &  &  &  & \\
		& Not eating any food & 7655 & 31.2 & 0.129 & 0.450 & 0.320\\
		& Carbohydrate $<$ 50\%\textsuperscript{*} & 4500 & 18.4 & 0.130 & 0.267 & 0.128\\
		& Carbohydrate $\geqslant$ 50\%\textsuperscript{\dag} & 12328 & 50.4 & 0.741 & 0.283 & 0.552\\
		9 am – 12 noon &  &  &  &  &  & \\
		& Not eating any food & 5447 & 22.3 & 0.237 & 0.079 & 0.401\\
		& Carbohydrate $<$ 50\% & 7227 & 29.5 & 0.158 & 0.492 & 0.173\\
		& Carbohydrate $\geqslant$ 50\% & 11809 & 48.2 & 0.605 & 0.429 & 0.426\\
		12 noon – 2 pm &  &  &  &  &  & \\
		& Not eating any food & 4783 & 19.5 & 0.156 & 0.356 & 0.019\\
		& Carbohydrate $<$ 50\% & 11112 & 45.4 & 0.405 & 0.413 & 0.560\\
		& Carbohydrate $\geqslant$ 50\% & 8588 & 35.1 & 0.439 & 0.231 & 0.421\\
		2 pm – 5 pm &  &  &  &  &  & \\
		& Not eating any food & 6926 & 28.3 & 0.130 & 0.123 & 0.659\\
		& Carbohydrate $<$ 50\% & 8277 & 33.8 & 0.249 & 0.602 & 0.076\\
		& Carbohydrate $\geqslant$ 50\% & 9280 & 37.9 & 0.621 & 0.276 & 0.266\\
		5 pm – 8 pm &  &  &  &  &  & \\
		& Not eating any food & 3043 & 12.4 & 0.114 & 0.199 & 0.034\\
		& Carbohydrate $<$ 50\% & 14240 & 58.2 & 0.516 & 0.590 & 0.639\\
		& Carbohydrate $\geqslant$ 50\% & 7200 & 29.4 & 0.370 & 0.211 & 0.328\\
		8 pm – 10 pm &  &  &  &  &  & \\
		& Not eating any food & 8722 & 35.6 & 0.322 & 0.291 & 0.480\\
		& Carbohydrate $<$ 50\% & 8898 & 36.4 & 0.266 & 0.551 & 0.212\\
		& Carbohydrate $\geqslant$ 50\% & 6863 & 28.0 & 0.412 & 0.158 & 0.308\\
		10 pm – 6 am &  &  &  &  &  & \\
		& Not eating any food & 16295 & 66.6 & 0.680 & 0.590 & 0.751\\
		& Carbohydrate $<$ 50\% & 4144 & 16.9 & 0.074 & 0.294 & 0.101\\
		& Carbohydrate $\geqslant$ 50\% & 4044 & 16.5 & 0.246 & 0.115 & 0.148\\
		\bottomrule
		\multicolumn{7}{l}{{\scriptsize \textit{Note: }}}\\
		\multicolumn{7}{l}{\scriptsize \textbf{Abbreviation:} LCA, latent class analysis, carb is short for carbohydrates.}\\
		\multicolumn{7}{l}{{\scriptsize \textsuperscript{*} Carbohydrate $<$ 50\% indicates that within the time slot, carbohydrate contributed $<$ 50\% total energy intake.}}\\
		\multicolumn{7}{l}{{\scriptsize \textsuperscript{\dag} Carbohydrate $\geqslant$ 50\% indicates that within the time slot, carbohydrate contributed $\geqslant$ 50\% total energy intake.}}\\
		\multicolumn{7}{l}{}\\ 
	\end{tabular}
\end{table}

@@

From the results of blood tests, 6.9\% of low carbohydrate eaters were found to be diabetic (diagnosed by HbA1C > 6.5\%), while the percentages of diabetes in the moderate and high carbohydrate eaters were 3.5\%, and 4.1\% (\textit{p} = 0.011), respectively. Although there was some evidence (\textit{p} = 0.027) that fasting blood glucose level may be slightly higher in non-diabetic low carbohydrate eaters, the geometric mean for HbA1C was probably lower (\textit{p} = 0.010) in moderate carbohydrate eaters (5.43, 95\%CI: 5.39, 5.47). Total cholesterol, HDL, and LDL were all lower in the moderate carbohydrate eaters, while no evidence of any difference of TG was found across three types of carbohydrate eaters.

@@

		\multicolumn{5}{l}{\textit{Note: }}\\

@@

		\multicolumn{5}{l}{{\scriptsize BMI body mass index; WC, waist circumference; HbA1C, haemoglobin HbA1C; DM, diabetes mellitus; TC, total }}\\
		\multicolumn{5}{l}{{\scriptsize cholesterol; HDL, high density lipoproteins; LDL, low density lipoproteins; TG, triglycerides.}} \\
		\multicolumn{5}{l}{{\scriptsize Glucose, HbA1C, TC, HDL, LDL, TG, and physical activity were expressed in geometric means (95\% CI) because the }} \\
		\multicolumn{5}{l}{ {\scriptsize data were positively skewed.}} \\

### 2018-09-04 paper/Chapters/Chapter4.tex
right before submission checked once more

However, when the MLCA extended the model to an individual level, three types of persons were further defined depending on their 4-day-diary: 1) low carbohydrate eaters, who mostly followed "regular meals day"; 2) moderate carbohydrate eaters, who had similar probabilities of following either "regular meals day" or "high percentage carbohydrate day"; 3) high carbohydrate eaters, who followed "high percentage carbohydrate day" for half of their survey. For the first time, as far as we know, the day-to-day food intake pattern variation within individuals was successfully captured by MLCA models. Results from the MLCA models showed that from the perspective of carbohydrate consumption, people were indeed changing their diet from day to day even over the captured four days. The MLCA models allowed the probability of following a certain type of carbohydrate eating day to vary across individuals. This properly accounted for the fact that for some people, their probability of following a type of food eating pattern during the survey could be higher/lower than that in the others. This finding also suggested that assuming a person will always follow a certain type of eating pattern is not appropriate.

@@

Third, detailed information on the occupations of the participants was not available, and thus we could not account for those who had a shift or night work. But considering that completing a 4-consecutive-day food diary may be a tremendous burden for all participants, those with shift or night work might be less able to comply and complete their diary for at 4 days. However, the latter hypothesis cannot be verified in the NDNS RP dataset. If people on shifts were included in the NDNS RP sample, the classification in both levels, as well as their associations with either hypertension or BMI would be biased.

Furthermore, under-reporting due to the burden of recording food and drink intake for 4 days for each occasion is also a concern in the current survey. To evaluate the influence of under-reporting in the NDNS RP sample, the study team conducted a doubly labelled water (DLW) sub-study within the survey. The details of this sub-study can be accessed from the Appendix X of the Official Reports provided by Public Health England \parencite{bates2014national,roberts2018national,NDNSofficial}. In the sub-study, there is an assumption that in healthy adults, if energy consumed in food matches total energy expended over a given time period, the individuals are deemed to be in energy balance. Using the DLW technique, total energy expenditure (TEE) from a sub-sample of NDNS RP participants were measured and compared with the estimates of energy consumption from their 4 day diaries. Results of the sub-study showed reasonable agreement between energy consumed and TEE (overall ratio: 0.73), indicating some potential misreporting. Reasons such as misreporting of actual consumption, under-reporting or modified usual intake due to the burden of the survey/DLW sub-study may contribute to under-reporting of their energy consumed. We cannot extrapolate the estimated under-reporting to the whole sample since other individuals' diet might be differentially affected by under-reporting. Moreover, within NDNS RP, it is not possible to differentiate between under-reporting due to ill health vs actual misreporting. Subsequently, most previous studies using data from the NDNS RP do not adjust for under-reporting of nutrients intake in the models.

@@

We have successfully defined carbohydrate eating patterns in the general population in UK adults using the NDNS RP database in both observation day level and individual level. Low carbohydrate eaters tended to have higher energy intake from fat and alcohol compared to other types of carbohydrate eaters. Moderate carbohydrate eaters reported the lowest total daily energy intake and tended to have a higher energy intake later in the day. High carbohydrate eaters obtained most of their carbohydrate as well as energy earlier in the day. These eating patterns specifically for carbohydrate intake were found to differ by timing, quantity, and contributions to energy consumption. Compared with low carbohydrate eaters, men with a moderate carbohydrate eating pattern had a lower prevalence of hypertension. Women in moderate carbohydrate eating pattern and lived alone had a larger waist circumference compared with low carbohydrate eaters. Among women who lived with partners, a high carbohydrate eating pattern was associated with both lower BMI and smaller waist circumferences. Longitudinal studies are needed to investigate whether the identified eating patterns themselves are changing over time and to study how such circadian eating patterns may relate to the change of blood pressure, obesity, and other health outcomes (incidence of cancer, cardiovascular diseases, or mortality).

>>>>

However, when the MLCA extended the model to an individual level, three types of persons were further defined depending on their 4-day-diary: 1) low carbohydrate eaters, who mostly followed "regular meals day"; 2) moderate carbohydrate eaters, who had similar probabilities of following either "regular meals day" or "high percentage carbohydrate day"; 3) high carbohydrate eaters, who followed "high percentage carbohydrate day" for half of their survey. For the first time, as far as we know, the day-to-day food intake pattern variation within individuals was successfully captured by MLCA models. Results from the MLCA models showed that from the perspective of carbohydrate consumption, people were indeed changing their diet from day to day even over the captured four days. The MLCA models allowed the probability of following a certain type of carbohydrate eating day to vary across individuals. This properly accounted for the fact that for some people, their probability of following a type of food eating pattern during the survey could be higher/lower than that in the others. This finding also suggested that assuming a person will always follow a certain type of temporal eating pattern is not appropriate.

@@

Third, detailed information on the occupations of the participants was not available, and thus we could not account for those who had a shift or night work. But considering that completing a 4-consecutive-day food diary may be a tremendous burden for all participants, those with shift or night work might be less able to comply and complete their diary for 4 days. However, the latter hypothesis cannot be verified in the NDNS RP dataset. If people on shifts were included in the NDNS RP sample, the classification in both levels, as well as their associations with either hypertension or BMI would be biased.

Furthermore, under-reporting due to the burden of recording food and drink intake for 4 days at each occasion is also a concern in the current survey. To evaluate the influence of under-reporting in the NDNS RP sample, the study team conducted a doubly labelled water (DLW) sub-study within the survey. The details of this sub-study can be accessed from the Appendix X of the Official Reports provided by Public Health England \parencite{bates2014national,roberts2018national,NDNSofficial}. In the sub-study, there is an assumption that in healthy adults, if energy consumed in food matches total energy expended over a given time period, the individuals are deemed to be in energy balance. Using the DLW technique, total energy expenditure (TEE) from a sub-sample of NDNS RP participants were measured and compared with the estimates of energy consumption from their 4 day diaries. Results of the sub-study showed reasonable agreement between energy consumed and TEE (overall ratio: 0.73), indicating some potential misreporting. Reasons such as misreporting of actual consumption, under-reporting or modified usual intake due to the burden of the survey/DLW sub-study may contribute to under-reporting of their energy consumed. We cannot extrapolate the estimated under-reporting to the whole sample since other individuals' diet might be differentially affected by under-reporting. Besides, within NDNS RP, it is not possible to differentiate between under-reporting due to ill health vs actual misreporting. Subsequently, most previous studies using data from the NDNS RP do not adjust for under-reporting of nutrients intake in the models.

@@

We have successfully defined carbohydrate eating patterns in the general population in UK adults using the NDNS RP database in both observation day level and individual level. Low carbohydrate eaters tended to have higher energy intake from fat and alcohol compared to other types of carbohydrate eaters and eating later in the day. Moderate carbohydrate eaters reported the lowest total daily energy intake and tended to have a higher energy intake later in the day. High carbohydrate eaters obtained most of their carbohydrate as well as energy earlier in the day. These eating patterns specifically for carbohydrate intake were found to differ by timing, quantity, and contributions to energy consumption. Compared with low carbohydrate eaters, men with a moderate carbohydrate eating pattern may had a lower prevalence of hypertension. Women in moderate carbohydrate eating pattern and lived alone had a larger waist circumference compared with low carbohydrate eaters. Among women who lived with partners, a high carbohydrate eating pattern was associated with both lower BMI and smaller waist circumferences. Longitudinal studies are needed to investigate whether the identified eating patterns themselves are changing over time and to study how such circadian eating patterns may relate to the change of blood pressure, obesity, and other health outcomes (incidence of cancer, cardiovascular diseases, or mortality).
